---
title: "Rolling Sales Queens"
author: "Benjamin Reddy/Yao Steven Jason"
date: "February 21, 2017"
output:
  word_document: default
  pdf_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#require(gdata)
#require(plyr) #Added by Monnie McGee
#install the gdata and plyr packages and load in to R.
setwd("C:\\Users\\Yao\\Dropbox")
```

So, save the file as a csv and use read.csv instead
```{r}
qns <- read.csv("rollingsales_queens.csv",skip=4,header=TRUE)
```

Check the data
```{r}
head(qns)
summary(qns)
str(qns) # Very handy function!
```

clean/format the data with regular expressions
More on these later. For now, know that the
pattern "[^[:digit:]]" refers to members of the variable name that
start with digits. We use the gsub command to replace them with a blank space.
We create a new variable that is a "clean' version of sale.price.
And sale.price.n is numeric, not a factor.

```{r, message=FALSE}
library(gdata)
library(plyr)
qns$SALE.PRICE.N <- as.numeric(gsub("[^[:digit:]]","", qns$SALE.PRICE))
count(is.na(qns$SALE.PRICE.N))
names(qns) <- tolower(names(qns)) # make all variable names lower case
## Get rid of leading digits
qns$gross.sqft <- as.numeric(gsub("[^[:digit:]]","", qns$gross.square.feet))
qns$land.sqft <- as.numeric(gsub("[^[:digit:]]","", qns$land.square.feet))
qns$year.built <- as.numeric(as.character(qns$year.built))
```

do a bit of exploration to make sure there's not anything
weird going on with sale prices
```{r}
attach(qns)
hist(sale.price.n) 
detach(qns)
```

keep only the actual sales
```{r}
qns.sale <- qns[qns$sale.price.n!=0,]
plot(qns.sale$gross.sqft,qns.sale$sale.price.n)
plot(log10(qns.sale$gross.sqft),log10(qns.sale$sale.price.n))
```

for now, let's look at 1-, 2-, and 3-family homes
```{r}
qns.homes <- qns.sale[which(grepl("FAMILY",qns.sale$building.class.category)),]
dim(qns.homes)
plot(log10(qns.homes$gross.sqft),log10(qns.homes$sale.price.n))
summary(qns.homes[which(qns.homes$sale.price.n<100000),])
```

remove outliers that seem like they weren't actual sales
```{r}
qns.homes$outliers <- (log10(qns.homes$sale.price.n) <=5) + 0
qns.homes <- qns.homes[which(qns.homes$outliers==0),]
plot(log10(qns.homes$gross.sqft),log10(qns.homes$sale.price.n))
```
